An Efficient Minibatch Acceptance Test for Metropolis-Hastings

### 2016-09-22 nips_2016.tex
More changes to Section 4, I am understanding Johns original intentions better

\newcommand{\simiid}{\overset{\textrm{i.i.d.}}{\sim}}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}

@@

Here we derive a new test which incorporates the variability in
minibatch statistics as {\em a natural part of the test}. Because of
this, the amount of data required for each test is fixed with high
probability and in expected value. We use a Barker test function
\cite{Barker65} rather than a Metropolis test, which makes our test
naturally error tolerant. The idea of using a noise-tolerant test
using Barker's test function was suggested but not explored
empirically in \cite{Bardenet15} section 6.3. But the asymptotic test
statistic CDF and the Barker function are different, which leads to
fixed errors for the approach in \cite{Bardenet15}. Here we show that
the difference between the distributions can be corrected with an
additive random variable. This leads to a test which is fast, and
whose error can be made arbitrarily small.

Our test is applicable when the variance (over data samples) of the
log acceptance probability is small enough (less than 1). Its not
clear at first why this quantity should be so bounded. But we will see
that it is ``natural'' for well-specified models running
Metroplis-Hastings sampling with optimal proposals
\cite{OptimalScaling01} on a full dataset. When we reduce the amount
of data for the test, the variance goes up, and we have to reduce
it in one of several ways. Either:

@@

Its worth discussing at this point the typical goals of
M-H sampling on very large datasets.  By the Bernstein-von Mises
Theorem, the posterior distribution of the parameter $\theta$ for a
Bayesian inference task is asymptotically normal, and has variance
that scales inversely with the number of data samples $N$. This mode
is extremely sharp for large datasets, which may contain millions or
billions of samples. Simply sampling from this distribution is one
application, but an efficient proposal distribution
\cite{OptimalScaling01} has similar variance to the target
distribution and will diffuse to it extremely slowly from an
initialization value which is (likely to be) many standard deviations
away. If there are any other strong modes, it is very likely for the
sampler to find one of them and become trapped in it when run at the
normal distribution temperature ($T=1$). A common solution is to anneal
the sampler, running first at high temperature (scaling log
likelihoods by $1/T$) which flattens the likelihood landscape.  This is
turn reduces the variance of the log acceptance probability and allows
our acceptance test to be applied.

A second question concerns step size. Once we have fixed temperature,
our variance constraint implies that we have to trade-off proposal
step size $s$ and batch size $b$ ($b \propto p^2$). i.e. we can make
many small steps, or one large step, with a given batch of data. One
of the primary drivers of this work is our belief in the value of
small steps. For applications to neural networks or other models where
the posterior is multimodel, posterior inference is arguably a search
process. Covering the search space densely with small steps is much
more valuable than few sparse steps toward the nearest optimum. In
this mode, Metropolis-Hastings can be used in similar fashion to
Stochastic Gradient Descent. The goal in SGD is to make gradual
progress to a posterior mode with each step, taking small steps so
that the cumulative displacement has progressively lower variance. A
substantial part of the computational work of MCMC on massive datasets
will be similarly in reaching a stationary distribution, which really
means finding a deep posterior mode. Taking noisy small steps will
nevertheless make steady progress to a posterior mode since their bias
is in that direction. We demonstrate this behavior in our logistic regression
experiments.

@@

Our goal is to sample from the posterior $p(\theta \mid x_1,\ldots,x_N)$, which
is a normal distribution $\mathcal{N}(\mu, 1/N)$ centered on the sample mean
$\mu$, and with variance $1/N$. In one dimension, an efficient proposal
distribution has the same variance as the target
distribution~\cite{OptimalScaling01}, so we use the proposal
$\mathcal{N}(\theta'-\theta,1/N)$, which is implemented as
$q(\theta'\mid\theta)=\phi((\theta'-\theta)N)$, where $\phi(x)$ is the Normal
density function with zero mean and unit variance. This proposal is symmetric
$q(\theta'\mid\theta)=q(\theta\mid\theta')$, and since we assumed a uniform
prior on $\theta$, we see that $\psi(u,\theta',\theta)$ reduces to $\log u$. Our
worst-case scenario is specified in Lemma~\ref{lem:worst_case}.
{\color{blue} Daniel: this might be a silly question, but why do we assume the
proposal has mean $\theta'-\theta$? I thought it would have a mean of $\theta$
because that is the current sample and the Gaussian is centered at that point?}

@@

Similar results can be shown for other distributions and proposals by
identifying regions in product space $(\theta,\theta'-\theta,u)$ such that the
hypothesis test needs to separate nearly-equal values.  It follows that the
accelerated M-H tests in~\cite{cutting_mh_2014,icml2014c1_bardenet14} require at
least a constant fraction $\geq c$ in the amount of data consumed per test
compared to full-dataset tests, so their speed-up is at most $1/c$.

@@

{\color{blue} Daniel: cleaned up this section, deleted redundant delta equation,
deleted a duplicate equation in Lemma~\ref{lem:detailed_balance}.}

@@

\subsection{Barker (Logistic) Acceptance Function}\label{ssec:barker_function}
For our new MH test we use the Barker logistic function:
$g(s)=(1+\exp(-s))^{-1}$ \cite{Barker65}. Straightforward arithmetic shows that it
satisfies the condition in Lemma~\ref{lem:detailed_balance}.  While it
is slightly less efficient than the Metropolis test when used on the
full dataset, we will see soon that its smoothness allows it to
naturally tolerate substantial variance in its input argument. This in
turn will lead to a much more efficient test on subsets of data.

Assume we begin with the current sample $\theta$  and a candidate sample
$\theta'$. Let $V$ be a uniform random variable $V \sim \mathcal{U}(0,1)$. We
accept $\theta'$ if $g(\Delta(\theta,\theta')) > V$, and reject otherwise.
Since $g(s)$ is monotonically increasing, its inverse $g^{-1}(s)$ is
well-defined and unique. So an equivalent test is to accept $\theta'$ iff
\begin{equation}\label{eq:equivalent_test}
    \Delta(\theta,\theta') > X = g^{-1}(V)
\end{equation}
where $X$ is a random variable with the logistic distribution (its CDF is the
logistic function). To see this notice that $\frac{dV}{dX} = g'$, that $g'$ is
the density corresponding to a logistic CDF, and finally that $\frac{dV}{dX}$ is
the density of $X$. The density of $X$ is symmetric, so we can equivalently test
whether
\begin{equation}\label{eq:the_exact_test}
    \Delta(\theta,\theta') + X > 0
\end{equation}
for a logistic random variable $X$. {\color{blue} Daniel: I think the text
before Equation~\ref{eq:equivalent_test} should be similar to what we had in
our NIPS submission because I find it much clearer.}

@@

We now move to acceptance testing using the minibatch estimator
$\Delta^*(\theta,\theta')$. From equation (\ref{eq:delta3}),
$\Delta^*(\theta,\theta')$ can be represented as a constant term plus the mean
of $b$ IID terms $\Lambda_i(\theta,\theta')$ of the form
$N\log\frac{p(x_i|\theta')}{p(x_i|\theta)}$. As $b$ increases,
$\Delta^*(\theta,\theta')$ therefore has a distribution which approaches a
normal distribution by the central limit theorem. Later we will provide specific
bounds between the CDF of $\Delta^*(\theta,\theta')$ and a Normal CDF, but for
now we use an asymptotic argument.

In the limit, since $\Delta^*$ is normally distributed about its mean $\Delta$,
we can write
\begin{equation}\label{eq:relationship}
    \Delta^* = \Delta + X_{\rm norm}, \quad X_{\rm norm} \sim \mathcal{\bar{N}}(0, \sigma^2(\Delta^*)),
\end{equation}
where $\mathcal{\bar{N}}(0, \sigma^2(\Delta^*))$ denotes a distribution which is
approximately normal with variance $\sigma^2(\Delta^*)$.  But to perform the
test in Equation~\ref{eq:the_exact_test} we want $\Delta + X$ for a logistic
random $X$ (call it $X_{\rm log}$ from now on). In \cite{Bardenet15} it was
proposed to use $\Delta^*$ in a Barker test anyway and tolerate the fixed error
caused by this approximation.

Our approach is to instead decompose $X_{\rm log}$ as
\begin{equation}\label{eq:deconvolution}
    X_{\rm log} = X_{\rm norm}+X_{\rm corr},
\end{equation}
where we assume $X_{\rm norm}$ has zero-mean, Normal distribution with
variance $\sigma^2$, and $X_{\rm corr}$ is a zero-mean ``correction''
variable with density $C_{\sigma}(X)$.  The two variables are added
(i.e., their distributions convolve) to form $X_{\rm log}$. It will be
convenient to express this relationship in terms of the CDFs $N_{\sigma}(X_{\rm
  norm}) = \Phi(X_{\rm norm}/\sigma)$ and $S(X_{\rm log})$, where
$\Phi(.)$ is the Normal CDF with zero mean and unit variance and
$S(.)$ is the standard logistic function.

Equation (\ref{eq:deconvolution}) implies that $S = N_{\sigma} * C_{\sigma}$ where $*$ denotes convolution. We will later
derive a numerical representation of $C$ by deconvolution, and using this we can generate
samples of $X_{\rm corr}$.

The acceptance test is now:
\begin{equation}\label{eq:criteria}
    \Delta + X_{\rm log} = (\Delta + X_{\rm norm}) + X_{\rm corr} = \Delta^* + X_{\rm corr} >0.
\end{equation}
i.e. given an estimate $\Delta^*$ from the current data minibatch (assuming its
variance is small enough) we test acceptance by adding a random variable $X_{\rm corr}$ and
then accept $\theta'$ if the result is positive.

If $\mathcal{\bar{N}}(0, \sigma^2(\Delta^*))$ is exactly $\mathcal{N}(0, \sigma^2(\Delta^*))$,
the above test is exact as well. And as we will see later,  if there is a maximum error $\epsilon$ between
the CDF of $\mathcal{\bar{N}}(0, \sigma^2(\Delta^*))$ and the CDF of $\mathcal{N}(0, \sigma^2(\Delta^*))$,
then the acceptance test has an error of at most $\epsilon$.

\begin{figure}[t]
    \centering
    \includegraphics[width=1\textwidth]{mh_convolution_diagram_v2}
    \caption{
    Three examples of $X_{\rm norm}$ and $X_{\rm corr}$ distributions that convolve to form the
    standard logistic distribution. We use three standard deviation values of $X_{\rm norm}$. The
    two red curves convolve to form the logistic, etc. The $y$-axis is capped at 0.5 for
    readability. This figure must be viewed in color.
    }
    \label{fig:deconvolution}
    %\vspace{-10pt}
\end{figure}

%% It is important to understand why we use Equations~\ref{eq:deconvolution} and~\ref{eq:criteria}. The
%% deconvolution process allows us to use $\Delta' + X_{\rm corr}$ as a proxy for $\Delta + X$. It
%% works because if we know the distribution\footnote{Computing the distributions of $\varepsilon$ and
%% $X_{\rm norm}$ reduces to finding their standard deviations.} of $\varepsilon$ (from
%% Equation~\ref{eq:relationship}), then we can pretend that our $\Delta'$ is in fact $\Delta$. To
%% ``insert'' $\varepsilon$ into our test, we deconvolve $X$ so that one of its components, $X_{\rm
%% norm}$, plays the part of $\varepsilon$. In light of our revised MH test, the previous discussion
%% raises the question: in an iteration of MCMC with current $\theta$ and proposed $\theta'$, how do
%% we know the distribution of $X_{\rm corr}$? We need the correct distribution to sample the
%% \emph{values} of $X_{\rm corr}$ appropriately. This procedure consists of three major steps.

%% \begin{enumerate}[noitemsep]
%%     \item We first need to estimate the standard deviation of $X_{\rm norm}$ (i.e. the standard deviation of $\Delta'$). We use one minibatch data to complete this task. From the proof of Lemma~\ref{lem:gaussian}, we know the randomness comes from $\frac{N}{n}\sum_{i=1}^n \left(\log p(x_i|\theta') - \log p(x_i|\theta) \right)$. Thus, ${\rm var}(\Delta') = \frac{N^2}{n^2}\sum_{i=1}^n{\rm var}\left (  \log p(x_i|\theta') - \log p(x_i|\theta) \right ) = \frac{N^2}{n}{\rm var}\left (  \log p(x_i|\theta') - \log p(x_i|\theta) \right )$. We use data instances in one minibatch to estimate the ${\rm var} \left (  \log p(x_i|\theta') - \log p(x_i|\theta) \right )$. Since there are $n$ data instances in one minibatch, the estimation of ${\rm var} \left (  \log p(x_i|\theta') - \log p(x_i|\theta) \right )$ has high accuracy. In our experiments (see Section~\ref{sec:experiments}) we also use the complete data to obtain the population variance of $\left (  \log p(x_i|\theta') - \log p(x_i|\theta) \right )$, and use the nonparametric Bootstrap to estimate the ${\rm std(std}(\Delta'))$, so that we can construct the confidence interval of ${\rm std}(\Delta')$ to reveal the accuracy of our estimation.

%% 	\item If the ${\rm std}(\Delta') > 1.2$, the decomposition of $X$ (i.e. Equation~\ref{eq:deconvolution}) is unstable. We consume more minibatches to decrease the ${\rm std}(\Delta')$. We update the value of $\Delta'$ by $\Delta' = (k\Delta'_{\rm old} + \Delta'_{\rm new}) / (k+1)$, where $\Delta'_{\rm old}$ is the existing $\Delta'$ which uses $k$ minibatches to computed (usually $k=1$), and $\Delta'_{\rm new}$ is the $\Delta'$ computed from the new minibatch. We also update the value of ${\rm std}(\Delta')$ by ${\rm std}(\Delta') = \sqrt{\left ( k^2 {\rm var}(\Delta'_{\rm old}) + {\rm var}(\Delta'_{\rm new}) \right ) / (k+1)^2}$. Thus, we can decrease the ${\rm std}(\Delta')$ by consuming more data until the ${\rm std}(\Delta') > 1.2$. With help of temperature (see Section~\ref{ssec:discussion}), we will show that usually the mean of minibatch consumed per iteration is below 2 in Section~\ref{sec:experiments}. Later our efficiency analysis in Section~\ref{sec:theory} will reveal that the number of data instance consumed by our algorithm is much fewer than that of existing algorithms.

%%     \item Once we have ${\rm std}(\Delta')$, we can determine the distribution of $X_{\rm corr}$
%%     from standard deconvolution techniques. These rely on the well-known fact that the Fourier
%%     transform of a convolution is the product of the individual Fourier transforms. In practice, we
%%     pre-compute distributions of $X_{\rm corr}$ for different ${\rm std}(\Delta')$ values to
%%     facilitate the sampling process.
%% \end{enumerate}

@@

Our goal is to compute the distribution of the correction variable $X_{\rm corr}$ such that
$X_{\rm norm} + X_{\rm corr} = X_{\rm log}$, where $X_{\rm norm} \in \mathcal{N}(0,\sigma)$
and $X_{\rm log}$ has a standard logistic CDF $(1+\exp(-X))^{-1}$.
In the next section we show that the accuracy of the test depends on the absolute
error between the CDFs of $X_{\rm norm} + X_{\rm corr}$ and the distribution of $X_{\rm log}$.
So we minimize this in our construction.

Our goal then is to minimize
\begin{equation}
  |\Phi_{\sigma} * C_{\sigma} - S|
\end{equation}
where $\Phi_{\sigma}(X) = \Phi(X/\sigma)$ and $\Phi$ is the standard
normal CDF, $S(X)$ is the logistic function, $C_{\sigma}(X)$ is the
{\em density} of the correction distribution, and $*$ denotes
convolution. In continuous form, this value is
\begin{equation}
  \label{eq:conv1}
  \sup_X\left|\int_{-V}^{V}\Phi_{\sigma}(X-Y) C_{\sigma}(Y) dY - S(X)\right|
  \end{equation}
Assuming $X_{\rm corr}$ lies in the interval $[-V,V]$. For computation of $C_{\sigma}(Y)$, we
discretize $X$ and $Y$ into $4N+1$ and $2N+1$ values respectively, with $X \in [-2V,2V]$ and
$Y \in [-V,V]$. Writing $X_i = i * V / N$ and $Y_j = j * V / N$, the condition in
equation (\ref{eq:conv1}) can be written:
\begin{equation}
  \max_{i \in \{-2N,\ldots,2N\}}\left|\sum_{j = -N}^{N}\Phi_{\sigma}(X_i-Y_j) C_{\sigma}(Y_j) - S(X_i)\right|
\end{equation}
and defining a matrix $M_{ij} = \Phi_{\sigma}(X_i-Y_j)$, vectors $U_j = C_{\sigma}(Y_j)$ and $V_i = S(X_i)$, the
above is equivalent to minimizing the 1-norm:
\begin{equation}
  \label{eq:error_norm}
  ||MU-V||_1
\end{equation}
wrt $U$. We have an additional constraint that $U_j > 0$ for all $j$, since $U$ represents
a density. We first explored optimizing this system with linear programming to find the $U$ such that:
\begin{equation}
  \begin{split}
    & \min(\epsilon) \qquad \hbox{where} \\
    & -\epsilon \leq MU - V \leq \epsilon \\
    & U \geq 0
  \end{split}
\end{equation}
which was tractable up to a few hundred dimensions. However, the discretization error is bounded by
the curvature of the underlying functions which are here slightly less than one. i.e.
the errors are of the order of $(V/N)^2$. Here we chose $V=20$ to provide adequate containment of
the distributions (the CDFs are extremely close to either zero or one outside this range). So
with 200 points, we have a discretization error of approximately 0.01. This is quite high. To yield
higher resolution and lower error, we switched to a least squares solution. 
Thus we seek to find $U$ that minimizes
\begin{equation}
  ||MU-V||_2
\end{equation}
Its not clear that this will yield a good 1-norm for equation (\ref{eq:error_norm}), but in practice it does.
Because the solution of this system is not very stable, especially in high dimensions, we actually
minimize a regularized version
\begin{equation}
  ||MU-V||_2 + \lambda||U||_2
\end{equation}
the solution is
\begin{equation}
  U = (M^T M + \lambda I)^{-1}V
\end{equation}
with this approach, there is no guarantee that $U \geq 0$. However, we have some
flexibility in the choice of $\sigma$ in equation (\ref{eq:conv1}). As we decrease
the variance of $X_{\rm norm}$, the variance of $X_{\rm corr}$ grows by the same amount
and is in fact the result of convolution with a gaussian whose variance is the difference.
Thus as $\sigma$ decreases, $C(X)$ grows and approaches the derivative of a
logistic function at $\sigma = 0$. It retains some very weak negative values for $\sigma > 0$
but removal of those values leads to very small error.

>>>>

\newcommand{\simiid}{\overset{\textrm{i.i.d.}}{\sim}}
\DeclareMathOperator*{\argmin}{arg\,min}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}

@@

Here we derive a new test which incorporates the variability in minibatch
statistics as {\em a natural part of the test}. Because of this, the amount of
data required for each test is fixed with high probability and in expected
value. {\color{blue} Daniel: we really need to show the amount of data is fixed
with ``high probability'' (the results suggest ``OK probability'').} We use a
Barker test function~\cite{Barker65} rather than a Metropolis test, which makes
our test naturally error tolerant. The idea of using a noise-tolerant test using
Barker's test function was suggested but not explored empirically
in~\cite{Bardenet15} section 6.3. But the asymptotic test statistic CDF and the
Barker function are different, which leads to fixed errors for the approach
in~\cite{Bardenet15}. Here we show that the difference between the distributions
can be corrected with an additive random variable. This leads to a test which is
fast, and whose error can be made arbitrarily small.

Our test is applicable when the variance (over data samples) of the log
acceptance probability is small enough (less than 1). Its not clear at first why
this quantity should be bounded, but we will show that it is ``natural'' for
well-specified models running Metropolis-Hastings sampling with optimal
proposals~\cite{OptimalScaling01} on a full dataset. When we reduce the amount
of data for the test, the variance goes up, and we have to reduce it in one of
several ways. Either:

@@

Its worth discussing at this point the typical goals of
M-H sampling on very large datasets.  By the Bernstein-von Mises
Theorem, the posterior distribution of the parameter $\theta$ for a
Bayesian inference task is asymptotically normal, and has variance
that scales inversely with the number of data samples $N$. This mode
is extremely sharp for large datasets, which may contain millions or
billions of samples. Simply sampling from this distribution is one
application, but an efficient proposal distribution~\cite{OptimalScaling01} has similar variance to the target
distribution and will diffuse to it extremely slowly from an
initialization value which is (likely to be) many standard deviations
away. If there are any other strong modes, it is very likely for the
sampler to find one of them and become trapped in it when run at the
normal distribution temperature ($T=1$). A common solution is to anneal
the sampler, running first at high temperature (scaling log
likelihoods by $1/T$) which flattens the likelihood landscape.  This is
turn reduces the variance of the log acceptance probability and allows
our acceptance test to be applied.

A second question concerns step size. Once we have fixed temperature,
our variance constraint implies that we have to trade-off proposal
step size $s$ and batch size $b$ ($b \propto p^2$), i.e. we can make
many small steps, or one large step, with a given batch of data. One
of the primary drivers of this work is our belief in the value of
small steps. For applications to neural networks or other models where
the posterior is multimodal, posterior inference is arguably a search
process. Covering the search space densely with small steps is much
more valuable than few sparse steps toward the nearest optimum. In
this mode, Metropolis-Hastings can be used in similar fashion to
Stochastic Gradient Descent. The goal in SGD is to make gradual
progress to a posterior mode with each step, taking small steps so
that the cumulative displacement has progressively lower variance. A
substantial part of the computational work of MCMC on massive datasets
will be similarly in reaching a stationary distribution, which really
means finding a deep posterior mode. Taking noisy small steps will
nevertheless make steady progress to a posterior mode since their bias
is in that direction. We demonstrate this behavior in our logistic regression
experiments.

@@

Our goal is to sample from the posterior $p(\theta \mid x_1,\ldots,x_N)$, which
is a normal distribution $\mathcal{N}(\mu, 1/N)$ centered on the sample mean
$\mu$, and with variance $1/N$. In one dimension, an efficient proposal
distribution has the same variance as the target
distribution~\cite{OptimalScaling01}, so we use the proposal
$\mathcal{N}(\theta'-\theta,1/N)$, which is implemented as
$q(\theta'\mid\theta)=\phi((\theta'-\theta)N)$, where $\phi(x)$ is the Normal
density function with zero mean and unit variance. This proposal is symmetric
$q(\theta'\mid\theta)=q(\theta\mid\theta')$, and since we assumed a uniform
prior on $\theta$, we see that $\psi(u,\theta',\theta)$ reduces to $\log u$. Our
worst-case scenario is specified in Lemma~\ref{lem:worst_case}.
{\color{blue} Daniel: this might be a silly question, but why do we assume the
proposal has mean $\theta'-\theta$? I thought it would have a mean of $\theta$
because that is the current sample and the Gaussian is centered at that point?
That will also clarify my confusion about the $\psi$ function.}

@@

Similar results can be shown for other distributions and proposals by
identifying regions in product space $(\theta,\theta'-\theta,u)$ such that the
hypothesis test needs to separate nearly-equal values.  It follows that the
accelerated M-H tests in~\cite{cutting_mh_2014,icml2014c1_bardenet14} require at
least a constant fraction $\geq c$ in the amount of data consumed per test
compared to full-dataset tests, so their speed-up is at most $1/c$.
{\color{blue} Daniel: doesn't our method also have this issue? In general, it
seems like any test that requires minibatch information to show deviation from
the full batch version will suffer from this problem. Why not just force us to
stop if we use too much data?}

@@

@@

\subsection{Barker (Logistic) Acceptance Function}\label{ssec:barker_function}
For our new MH test we use the Barker logistic~\cite{Barker65} function:
$g(s)=(1+\exp(-s))^{-1}$. Straightforward arithmetic shows that it satisfies the
condition in Lemma~\ref{lem:detailed_balance}.  While it is slightly less
efficient than the Metropolis test when used on the full dataset, we will show
that its smoothness allows it to naturally tolerate substantial variance in its
input argument. This in turn will lead to a much more efficient test on subsets
of data. {\color{blue} Daniel: aside from the experiments, where do we show it
``naturally tolerates substantial variance'' in the input?}

Assume we begin with the current sample $\theta$ and a candidate sample
$\theta'$, and that $V \sim \mathcal{U}(0,1)$ is a uniform random variable. We
accept $\theta'$ if $g(\Delta(\theta,\theta')) > V$, and reject otherwise.
Since $g(s)$ is monotonically increasing, its inverse $g^{-1}(s)$ is
well-defined and unique. So an equivalent test is to accept $\theta'$ iff
\begin{equation}\label{eq:equivalent_test}
    \Delta(\theta,\theta') > X = g^{-1}(V)
\end{equation}
where $X$ is a random variable with the logistic distribution (its CDF is the
logistic function). To see this notice that $\frac{dV}{dX} = g'$, that $g'$ is
the density corresponding to a logistic CDF, and finally that $\frac{dV}{dX}$ is
the density of $X$. The density of $X$ is symmetric, so we can equivalently test
whether
\begin{equation}\label{eq:the_exact_test}
    \Delta(\theta,\theta') + X > 0
\end{equation}
for a logistic random variable $X$. {\color{blue} Daniel: I think the text
before Equation~\ref{eq:equivalent_test} should be similar to what we had in
our NIPS submission because I find it much clearer.}

@@

We now describe acceptance testing using the minibatch estimator
$\Delta^*(\theta,\theta')$. From Equation~\ref{eq:delta3},
$\Delta^*(\theta,\theta')$ can be represented as a constant term plus the mean
of $b$ IID terms $\Lambda_i(\theta,\theta')$ of the form
$N\log\frac{p(x_i|\theta')}{p(x_i|\theta)}$. As $b$ increases,
$\Delta^*(\theta,\theta')$ therefore has a distribution which approaches a
normal distribution by the Central Limit Theorem. We now describe this using an
asymptotic argument and defer specific bounds between the CDFs of
$\Delta^*(\theta,\theta')$ and a Gaussian to Section~\ref{sec:analysis}.

In the limit, since $\Delta^*$ is normally distributed about its mean $\Delta$,
we can write
\begin{equation}\label{eq:relationship}
    \Delta^* = \Delta + X_{\rm norm}, \quad X_{\rm norm} \sim \mathcal{\bar{N}}(0, \sigma^2(\Delta^*)),
\end{equation}
where $\mathcal{\bar{N}}(0, \sigma^2(\Delta^*))$ denotes a distribution which is
approximately normal with variance $\sigma^2(\Delta^*)$.  But to perform the
test in Equation~\ref{eq:the_exact_test} we want $\Delta + X$ for a logistic
random variable $X$ (call it $X_{\rm log}$ from now on). In~\cite{Bardenet15} it
was proposed to use $\Delta^*$ in a Barker test anyway and tolerate the fixed
error caused by this approximation.

Our approach is to instead decompose $X_{\rm log}$ as
\begin{equation}\label{eq:deconvolution}
    X_{\rm log} = X_{\rm norm}+X_{\rm corr},
\end{equation}
where we assume $X_{\rm norm} \sim \mathcal{N}(0, \sigma^2)$ and $X_{\rm corr}$
is a zero-mean ``correction'' variable with density $C_{\sigma}(X)$.  The two
variables are added (i.e., their distributions convolve) to form $X_{\rm log}$.
This decomposition requires an appropriate $X_{\rm corr}$ distribution. We show
in Section~\ref{sec:correction} that it is possible to use deconvolution to
derive a highly accurate representation of the density $C_\sigma(X)$.
Figure~\ref{fig:deconvolution} shows three examples of $X_{\rm norm}$ and
$X_{\rm corr}$ distributions that sum to $X_{\rm log}$.

Using $X_{\rm corr}$ samples from $C_\sigma(X)$, the acceptance test is now
\begin{equation}\label{eq:criteria}
    \Delta + X_{\rm log} = (\Delta + X_{\rm norm}) + X_{\rm corr} = \Delta^* + X_{\rm corr} >0.
\end{equation}
Therefore, assuming the variance of $\Delta^*$ is small enough, if we have an
estimate of $\Delta^*$ from the current data minibatch, we test acceptance by
adding a random variable $X_{\rm corr}$ and then accept $\theta'$ if the result
is positive (and reject otherwise).

If $\mathcal{\bar{N}}(0, \sigma^2(\Delta^*))$ is exactly $\mathcal{N}(0,
\sigma^2(\Delta^*))$, the above test is exact as well, and as we show in
Section~\ref{sec:analysis}, if there is a maximum error $\epsilon$ between the
CDF of $\mathcal{\bar{N}}(0, \sigma^2(\Delta^*))$ and the CDF of $\mathcal{N}(0,
\sigma^2(\Delta^*))$, then the acceptance test has an error of at most
$\epsilon$ relative to the full batch version. {\color{blue} Daniel: this
subsection is improved from earlier. My main concern now is that $X_{\rm norm}$
is described as being both approximately Gaussian $\bar{\mathcal{N}}$ and
Gaussian $\mathcal{N}$ here. We may want to clear that up; in our NIPS
submission we actually used a new variable $\varepsilon$ for that and assigned
it to be the $\mathcal{N}$ distribution.}

\begin{figure}[t]
    \centering
    \includegraphics[width=1\textwidth]{mh_convolution_diagram_v2}
    \caption{
    Three examples of $X_{\rm norm}$ and $X_{\rm corr}$ distributions that
    convolve to form the standard logistic distribution. We use three standard
    deviation values of $X_{\rm norm}$. The two red curves convolve to form the
    logistic, etc. The $y$-axis is capped at 0.5 for readability. This figure
    must be viewed in color. {\color{blue} Daniel: TODO Change figure to use
    LaTeX for the titles.}
    }
    \label{fig:deconvolution}
    %\vspace{-10pt}
\end{figure}

@@

Our proposed test in Equation~\ref{eq:criteria} requires knowing the
distribution of the correction variable $X_{\rm corr}$ such that $X_{\rm norm} +
X_{\rm corr} = X_{\rm log}$, where $X_{\rm norm} \sim \mathcal{N}(0,\sigma^2)$
and $X_{\rm log}$ has a standard logistic CDF, $(1+\exp(-X))^{-1}$. In
Section~\ref{sec:analysis}, we show that the accuracy of the test depends on the
absolute error between the CDFs of $X_{\rm norm} + X_{\rm corr}$ and $X_{\rm
log}$. Consequently, we need to minimize this in our construction of $X_{\rm
corr}$. More formally, let $\Phi_\sigma(X) = \Phi(X/\sigma)$ where $\Phi$ is the
standard normal CDF\footnote{Hence, $\Phi_\sigma$ is the CDF of a zero-mean
Gaussian with variance $\sigma$.}, $S(X)$ be the logistic function, and
$C_{\sigma}(X)$ be the \emph{density} of the correction $X_{\rm corr}$
distribution. Our goal, based on Equation~\ref{eq:deconvolution}, is to solve
the following optimization problem:
\begin{equation}\label{eq:overall_corr_problem}
    C_\sigma^* = \argmin_{C_\sigma} |\Phi_{\sigma} * C_{\sigma} - S|
\end{equation}
where $*$ denotes convolution. {\color{blue} Daniel: This is very unusual use of
convolutions (i.e., CDF * DENSITY = CDF). It's valid but I was very surprised to
know that.} For computation of $C_\sigma$, we assume that its input $Y$ and
another variable $X$ lie in the intervals $[-V,V]$ and $[-2V,2V]$, respectively.
{\color{blue} Daniel: I am a little confused at the different interval sizes.}
In continuous form, the optimization problem is therefore
\begin{equation}\label{eq:conv1}
    C_\sigma^* = \argmin_{C\sigma} \; \sup_{x \in [-2V,2V]}\left|\int_{-V}^{V}\Phi_{\sigma}(x-y) C_{\sigma}(y)dy - S(x)\right|
\end{equation}
To get $C_\sigma^*$ in a practical form, we discretize $X$ and $Y$ into $4N+1$
and $2N+1$ values respectively. Writing $X_i = i*V/N$ and $Y_j = j*V/N$, the
condition in Equation~\ref{eq:conv1} can be re-expressed as:
\begin{equation}\label{eq:conv2}
    C_\sigma^* = \argmin_{C_\sigma} \max_{i \in \{-2N,\ldots,2N\}}\left|\sum_{j = -N}^{N}\Phi_{\sigma}(X_i-Y_j) C_{\sigma}(Y_j) - S(X_i)\right|.
\end{equation}
Defining a matrix $M$ and vectors $u$ and $v$ such that $M_{ij} =
\Phi_{\sigma}(X_i-Y_j)$, $u_j = C_{\sigma}(Y_j)$, and $v_i = S(X_i)$,
Equation~\ref{eq:conv2} is equivalent to minimizing the $L_1$ norm:
\begin{equation}\label{eq:optimization_l1}
    u^* = \argmin_{u} \|Mu-v\|_1.
\end{equation}
{\color{blue} Daniel: Equation~\ref{eq:optimization_l1} is finding
the $C_\sigma$ minimizing the $L_1$ norm, but the previous one,
Equation~\ref{eq:conv2}, is finding the $C_\sigma$ that is minimizing the
$L_\infty$ norm, i.e., the maximum value! If we want this to be the $L_1$ norm
(and I think we do) then Equation~\ref{eq:conv2} needs an outer sum over the $i$
terms. In addition, Equation~\ref{eq:conv1} also needs to be modified, perhaps
eliminate the ``sup'' but keep $x \in [-2V,2V]$}

We have the additional constraint that $u_j > 0$ for all $j$, since $u$
represents a density. We first explored optimizing this problem with linear
programming to find a $u$ minimizing $|Mu - v|$ subject to $u \ge 0$. This was
tractable up to a few hundred dimensions for $u$. However, the discretization
error is bounded by the curvature of the underlying functions which are here
slightly less than one, i.e.  the errors are of the order of $(V/N)^2$.
{\color{blue} Daniel: unfortunately, I do not understand what this means (i.e.,
the curvature and how the $(V/N)^2$ is obtained).} Here we chose $V=20$ to
provide adequate containment of the distributions (the CDFs are extremely close
to either zero or one outside this range). So with 200 points, we have a
discretization error of approximately 0.01, which is relatively large.  To yield
higher resolution and lower error, we switched to a least squares solution:
\begin{equation}\label{eq:optimization_l2}
    u^* = \arg_u \min \; \|Mu-v\|_2^2 + \lambda \|u\|_2^2,
\end{equation}
where we add the standard regularization factor $\lambda$ for stability in high
dimensions. The solution is well-known from the normal equations: $u^* = (M^TM +
\lambda I)^{-1}M^Tv$, and in practice, it yields an acceptable $L_1$ norm for
Equation~\ref{eq:optimization_l1}.

With this approach, there is no guarantee that $u^* \geq 0$. However, we have
some flexibility in the choice of $\sigma$ in Equation~\ref{eq:conv1}.
{\color{blue} Daniel: is this the correct equation?} As we decrease the variance
of $X_{\rm norm}$, the variance of $X_{\rm corr}$ grows by the same amount and
is in fact the result of convolution with a Gaussian whose variance is the
difference.  Thus as $\sigma$ decreases, $C(X)$ grows and approaches the
derivative of a logistic function at $\sigma = 0$. It retains some very weak
negative values for $\sigma > 0$ but removal of those values leads to very small
error.
